---
title: "Class_11"
author: "Zaida Rodriguez (PID:A59010549)"
date: "11/5/2021"
output:
 pdf_document:
  latex_engine: xelatex
---

```{r}
pdb.data <- read.csv("Data Export Summary.csv", row.names = 1)
#pdb.data
```

> Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

The Xray is 87.53% and Em is 4.95%
```{r}
library(skimr)
skim(pdb.data)
round(sum(pdb.data$X.ray) / sum(pdb.data$Total) * 100, 2)
round(sum(pdb.data$EM) / sum(pdb.data$Total) * 100, 2)

#  to see all the column row percentage: 
round((colSums(pdb.data) / sum(pdb.data$Total)) *100, 2)
```


> Q2: What proportion of structures in the PDB are protein?

The proportion is 0.87

```{r}
#call for protein only
prot <- pdb.data$Total[1]

#call for the proportion of protein 
prot.prop <- prot/sum(pdb.data$Total)
prot.prop
```

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There were 22,824 protease protein structures


## 2. Visualizing the HIV-1 Protease Structures 

See VMD Image
![(vmdscene.png)

> Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

We only see one because it can be overwhelming to see the all the atoms separate. Also we know that all 3 atoms are binded together so no need to see them separately

> Q5: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?

HOH 308




