---
title: "Class_19.pdf"
author: "Zaida Rodriguez (PID:A59010549)"
date: "12/1/2021"
output: pdf_document
---

# Section 1. Proportion on G/G in a population. 

Downloaded a CSV and read in

```{r}
mxl <- read.csv("SampleGenotypes.csv")
head(mxl)
```

```{r}
mxl$Genotype..forward.strand.
```
```{r}
table(mxl$Genotype..forward.strand.)
```

Q. what is the proportion of G|G

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) * 100
```
> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

Read in file
```{r}
expr <- read.table("sample_geno_exp.txt")
head(expr)
```

```{r}
nrow(expr)
table(expr$geno)
```

```{r}
library(ggplot2)
ggplot(expr) + 
  aes(geno, exp, fill=geno) +
  labs(x="genotype", y="expression")+
  geom_boxplot(notch=T) +
  geom_point(alpha=0.1)
```
>Q14. Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

A/A expression is higher compared to the G/G plot. Therefore, having the G/G in this location leads to reduced expression of this gene. This SNP effects the expression of ORMDL3

```{r}
gel <- boxplot(expr$exp~expr$geno, main="Gene Expression Levels", xlab="Genotype", ylab="Expression")
```
To find the median:
```{r}
head(gel)
```
We know in `$stats` the third row stands for the median, therefore, we can focus in on that
```{r}
gel$stats[3,]
```

