---
title: "Machine Learning 1"
author: "Zaida Rodriguez (PID:A59010549)"
date: "10/22/2021"
output: pdf_document
---

#Clustering methods

Kmeans clustering in R is done with the `kmeans()` function
Here we make up some data to test and learn with.
```{r}
tmp <- c(rnorm(30, 3), rnorm(30,-3))
data <- cbind(x=tmp, y=rev(tmp))

#the goal of this is to make a data set that has -3 and +3 values in x and y 
# x:(-3, +3) and y:(+3, -3)
plot(data)
```
Run `kmeans()` set k(centers) to 2 and nstart to 20. The thing with Kmeans is you have to tell it how many clusters you want. 

```{r}
km <- kmeans(data, centers=2, nstart=2)
km
```
Clustering vector is telling you for which cluster your element belongs to.

> Q1. How many points are in each cluster?

```{r}
km$size
```

>Q2. What 'component of your result object details cluster assignment/membership?

```{r}
km$cluster
```


>Q3. What 'component' of your result object details cluster center?

```{r}
km$centers
```


>Q4. Plot x colored by the kmeans cluster assignment and add clsuters as blue points (or by clusters)

```{r}
plot(data, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```


#Hierarchical clustering

We will use the `hclust()` function on the same data as before and see how this method works.

```{r}
hc <- hclust(dist(data))
hc

#this function requires you to tell it the distance from each point
```


hclust has a plot method 

```{r}
plot(hc)
abline(h=7, col="blue")
# the bottom (leaves) are your row names
# it puts what is closest to each other together
# 2 main groups; you can "cut it"
```


To find our membership vector we need to "cut" the tree (dendrogram) and for this we use the `cutree()` fucntion and tell it the height to cut at.

```{r}
cutree(hc,h=7)

```

We can also use t
we want ...
```{r}
grps <- cutree(hc, k=2)
```

```{r}
plot(data, col=grps)
```


# Principal Component Analysis (PCA)

## PCA OF UK Food data 

read the file ine
```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
# there are 17 rows and 5 columns 
```

view the first 6 rows
```{r}
head(x)
```

fix the row names:
```{r}
rownames(x) <-  x[,1]
x <- x[,-1]
x
```

> Q2. Which approach to solving the 'row-names problem' mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

Don't do things this way, you will overwrite your object. Instead, read it in with row names already(seen below). This is the preferred method as it will not mess with your data and is more robust.
```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names = 1)
x
```

now create a barplot. By adding "col=rainbow" you can change the color of the bars
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
# barplot need to be run as a matrix, since we didn't have a matrix, you can force it right into the function. 
```
> Q5. Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

This pairwise plot allows you to compare the 4 countries against each other. 
The first row compares England (x) and Wales (y), then Scotland (y), then N. Ireland(y)
The 2nd row compares Wales (x) and England(y), then N. Ireland (y); and so on and so on.
The points that are outliers are visible and those indicate differences in the consumption of the particular food categories depending on the country.
The points that lie on the diagonal for a given plot indicate that the values are the same for both countries that are being compared.
```{r}
mycols <- rainbow(nrow(x))
pairs(x, col=mycols, pch=16)
```
> Q6. What is the main differences in between N. Ireland and the other countries of the UK interms of this data-set?

N, Ireland has a greater variation of the food consumption in comparison to England, Wales, and Scotland. 


## PCA to the rescue!
Not easy to interpret, takes time. Instead try PCA!

Here we will use the base R function for PCA, which is called `prcomp()`.
This function wants you to first transpose your data (ie: flip the columns with rows)
```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
plot(pca)
```

We want a score plot (aka: PCA plot). Basically plot of PC1 vs PC2

```{r}
attributes(pca)
```

> Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
#plot PC1 vs PC2
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1:2], labels=colnames(x))
```

> Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at the start of this document. 

```{r}
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1:2], labels=colnames(x), col=rainbow(nrow(x)))
```

We can also examine the PCA "loading", which tells us how much of the original variables contribute to each new PC.

```{r}
barplot(pca$rotation[,1], las=2)
```
> Q9. Generate a similar 'loadings plot' for PC2. What 2 food gorups feature prominantely and what does PC2 mainly tell us about?

```{r}
barplot(pca$rotation[,2], las=2)
# the 2 main food groups are fresh potatoes and soft drinks. It tells us how much of the original variables contribute to each new PC.
```


## one more PCA 

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```
> Q10. How many genes are there in this data?

```{r}
nrow(rna.data)
# there are 100 genes
```


```{r}
ncol(rna.data)
```

```{r}
colnames(rna.data)
```

```{r}
pca.rna = prcomp(t(rna.data), scale=TRUE)
summary(pca.rna)

```

```{r}
plot(pca.rna)
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels=colnames(rna.data))
```
PC1 is telling us which genes are changing the most between KO to KO and WT to WT. 
PC2 