---
title: "Candy_data"
author: "Zaida Rodriguez (PID:A59010549)"
date: "10/29/2021"
output: github_document
---

```{r}
candy_file <- "candy-data.csv"
candy = read.csv("candy-data.csv", row.names=1)
head(candy)
#View(candy)
```

> Q1. How many different candy types are in this dataset?

85

```{r}
nrow(candy)
```


> Q2. How many fruity candy types are in the dataset?

38

```{r}
dim(candy)
sum(candy$fruity)
```

```{r}
candy["Twix", ]$winpercent
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

37.35%

```{r}
candy["Red vines", ]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

76.77%

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

49.65% 
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent is measured differently

> Q7. What do you think a zero and one represent for the candy$chocolate column?

True or false 

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

No

> Q10. Is the center of the distribution above or below 50%?

Below

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

chocolate is higher ranked

```{r}
as.logical(candy$chocolate)
candy[as.logical(candy$chocolate),]
chocolate <- candy[as.logical(candy$chocolate),]$winpercent
mean(chocolate)
```


```{r}
fruity <- candy[(as.logical(candy$fruity)),]$winpercent
mean(fruity)
```

> Q12. Is this difference statistically significant?

Yes because it is not equal, so it is significantly different

```{r}
t.test(chocolate,fruity)
```

#3. Candy rankings
Lets make a varplot of the winpercent values for the various candy types

> Q13. What are the five least liked candy types in this set?

Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters

```{r}
head(candy[order(candy$winpercent),], n=5)
# candy %>% arrange(winpercent) %>% head(5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?



```{r}

```

> Q15. Make a first barplot of candy ranking based on winpercent values

```{r}
library(ggplot2)

ggplot(candy) +
    aes(winpercent, rownames(candy))+
    geom_col()
```

> Q16. Improve this by re-ordering the candy by the winpercent value

```{r}
ggplot(candy) +
    aes(winpercent, reorder(rownames(candy), winpercent)) +
    geom_col()
```

add color
```{r}
#color vector all black to start
my_cols=rep("black", nrow(candy)) #nrow(candy) is a little vector that is the same lenght as candy file

# now overwrite the chocolate entries with "chocolate"
my_cols[as.logical(candy$chocolate)] = "chocolate" #as.logical (t/f) anything is true will be chocolate
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

print it out 
```{r}
my_cols
```

```{r}
ggplot(candy) +
    aes(winpercent, reorder(rownames(candy), winpercent)) +
    geom_col(fill=my_cols)
```

> Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Sixlets

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is Starburst

# 4. Pricepoint

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
To remove these weird labels 
```{r}
rownames(candy) <- gsub("Õ","'", rownames(candy))
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's Miniatures

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

Nik L Nip, Nestle Smarties, Ring pops, Hershey's Krackel, Hershey's Milk Chocolate

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head(candy[ord,c(11,12)], n=5)
```

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
this is a correlation structure. 
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity and chocolate

> Q23. Similarly, what two variables are most positively correlated?

winpercent and chocolate

#PCA

```{r}
pca <- prcomp(candy, scale = T)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```
change the colors
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```


```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

fruity
